### 2024-04-01 presentation/main.tex
latest panels

\documentclass{beamer}
%Information to be included in the title page:
\title{Elite individuals, insitutions, and economic growth accounting}
\author{James Xu}
\institute{ECON 442, Duke University}
\date{2024}

\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{Sample frame title}
This is some text in the first frame. This is some text in the first frame. This is some text in the first frame.
\end{frame}

\begin{frame}
    \frametitle{Sample frame title}
    This is some text in the first frame. This is some text in the first frame. This is some text in the first frame.
    \end{frame}

\end{document}